---
title: "Case Study 1"
output: html_document
---
This needs the libraries "reshape" and "ggplot2"  ,"MultiClust" from  install.packages("BiocManager") 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(plyr)
library(data.table)
library(ggthemes)
library(ggpubr)
#Read in the beers and breweries files
Beers = read.csv("data/Beers.csv", stringsAsFactors = FALSE)
Breweries = read.csv("data/Breweries.csv", stringsAsFactors = FALSE)

#Tidy up the data

#Trim whitespace from the State abbreviation

Breweries$State = trimws(Breweries$State)

# Set a common Brewery ID variable to merge on: 

names(Breweries)[1] = "Brewery_id"

```

<br/>

#### Introduction:
The Hoppy Specter Brewery is looking to expand into a new U.S. State with their world-renowned heavily hopped "Pale as a Ghost" pale ale. Given the overwhelming bitterness of both their ethos and beer, they need to find the best home for their new location. They've asked Unique Consulting Svcs., LLC. to help them identify the new location.

<br/>

#### Question 1

Given the big increase in local breweries over the last several years we decided to take a look at the number of breweries in each state. While we would not recomend opening in an overly saturated market, we still think a place that has an established local brewery scene would be best. We wrote some code that helps us to identify the number of breweries present in each state. 

```{r q1}
brew_count <- count(Breweries$State)
colnames(brew_count) <- c("State", "Brewery_Count")
brew_count <- brew_count[with(brew_count,order(-Brewery_Count)), ]
# head(brew_count, 10)
knitr::kable(brew_count)
```

<br/> 

As we can see from the above, Colorado's market is likely saturated. States like Pennsylvania, Massachussets, Washington or Indiana might be more ripe for new opportunities. 

<br/>

#### Preparing the data for deeper analysis: 

<br/>

In order to provide the best possible data for our customer, Unique Consulting drew upon multiple data sets to best inform their decision making. Unfortunately, this data required some work to put in a proper form to support analysis. No billable hours were spent on these tasks. Tasks included: 

1. Merging the data sets into a single data set capable of allowing comparisons needed. 

```{r q2 }
# Merge the two dataframes

MergedBrewing = merge(Beers,Breweries, by = "Brewery_id")

# Fix the names conflict

names(MergedBrewing)[2] = "Beer_Name"
names(MergedBrewing)[8] = "Brewery_Name"
head(MergedBrewing)


```

As an example, here are the first 6 rows: 

```{r}
knitr::kable(head(MergedBrewing,6))
```

and the last 6 rows: 
```{r}
knitr::kable(tail(MergedBrewing,6))
```

2. In order to prepare for this analysis, we also needed to identify the number of values missing in the data set. We found the following results: 

<br/>

```{r q3}

knitr::kable(colSums(is.na(MergedBrewing)))


```


<br/>  

At some points in the analysis below, the missing values are not considered. We've identified those areas of concern below. 

<br/>  

#### The market for "Pale as a Ghost" Pale Ale: 

The "Pale as a Ghost" pale ale is the flagship beer for The Hoppy Specter Brewery and it scores high in ABV (alchol by volume) and IBU (bitterness). We would like to get a broad overview of the general taste trends for each state when it comes to ABV and IBU. The plots below show the median ABV and IBU for each state.

```{r q4_ibu_medians, fig.height= 10}
#Get Medians
MergedBrewing <- na.omit(MergedBrewing)
state_ibu_medians <- tapply(MergedBrewing$IBU, MergedBrewing$State, median, na.rm = T)
state_ibu_medians <- data.frame(state_ibu_medians)

#Reset Column Names
setDT(state_ibu_medians , keep.rownames = TRUE)
colnames(state_ibu_medians) <- c("State", "IBU_Medians")

#Bar Plot
p <- ggplot(data = state_ibu_medians, aes(x=State, y=IBU_Medians, width = .5)) +
    labs(title="Median IBU", x="State", y = "IBU") 

p + geom_bar(stat="identity") +
    coord_flip() +
    theme_minimal() + 
    theme(plot.title = element_text(hjust = 0.5), legend.position = "none")
```


```{r q4_abv_medians, fig.height= 10}

#ABV Medians
state_abv_medians <- tapply(MergedBrewing$ABV, MergedBrewing$State, median, na.rm = T)
state_abv_medians <- data.frame(state_abv_medians)

#Reset Column Names
setDT(state_abv_medians , keep.rownames = TRUE)
colnames(state_abv_medians) <- c("State", "ABV_Medians")

#Bar Plot
p <- ggplot(data = state_abv_medians, aes(x=State, y=ABV_Medians, width = .5)) + 
    labs(title="Median ABV", x="State", y = "ABV")

p + geom_bar(stat="identity") +
    coord_flip() +
    theme_minimal() + 
    theme(plot.title = element_text(hjust = 0.5), legend.position = "none") 


```


<br/>

#### Question 5
Given that our client\'s beer is very strong and bitter, we would like to study markets with these same specs. The code below returns the states with the highest ABV and IBU.
```{r q5}
#State with max ABV
MergedBrewing[which.max(MergedBrewing$ABV), "State"]

#State with max IBU
MergedBrewing[which.max(MergedBrewing$IBU), "State"]
```

<br/>  

#### Question 6
```{r q6}
summary(MergedBrewing$ABV)
```

<br/>

#### Question 7
```{r q7}
ggscatter(MergedBrewing,x="ABV",y="IBU",add="reg.line",conf.int="T",cor.coef="T",color = "blue",ellipse.type="norm",ellipse="F")+theme_minimal()+ stat_density_2d(aes(fill = ..level..), geom = "polygon")+scale_fill_continuous(low="lavenderblush", high="purple")
```
