---
title: "Case Study 1"
output: html_document
---

```{r setup, echo = FALSE}

library(ggplot2)
library(plyr)
library(data.table)
library(ggthemes)

beers <- read.csv("./data/Beers.csv")
colnames(beers)[1] <- c("Beer_Name")

breweries <- read.csv("./data/Breweries.csv")
colnames(breweries)[2] <- c("Brewery_Name")
breweries$State <- trimws(breweries$State)
```

<br/>

#### Introduction:
The Hoppy Specter Brewery is looking to expand into a new U.S. State with their world-renowned heavily hopped "Pale as a Ghost" pale ale. Given the overwhelming bitterness of both their ethos and beer, they need to find the best home for their new location. They've asked Unique Consulting Svcs., LLC. to help them identify the new location.

<br/>

#### Question 1

Given the big increase in local breweries over the last several years we decided to take a look at the number of breweries in each state. While we would not recomend opening in an overly saturated market, we still think a place that has an established local brewery scene would be best. This code creates a dataframe of states and the number of breweries.

```{r q1}
brew_count <- count(breweries$State)
colnames(brew_count) <- c("State", "Brewery_Count")
brew_count <- brew_count[with(brew_count,order(-Brewery_Count)), ]
# head(brew_count, 10)
knitr::kable(brew_count)
```



<br/>

#### Question 2
This code merges the Beers and Breweries data on the brewery id (variable names Brewery_id and Brew ID respectivley).

```{r q2 }
final <- merge(beers, breweries, by.x=c("Brewery_id"), by.y=c("Brew_ID"))
knitr::kable(head(final, 6))
```


<br/>

#### Question 3
This code finds the NA values and then sums the totals for each column. This will allow us to see if there is any data we might need to impute.

```{r q3}
knitr::kable(colSums(is.na(final)))

```


<br/>

#### Question 4
The "Pale as a Ghost" pale ale is the flagship beer for The Hoppy Specter Brewery and it scores high in ABV (alchol by volume) and IBU (bitterness). We would like to get a broad overview of the genral taste trends for each state when it comes to ABV and IBU. The code below gets the median ABV and IBU for each state and then creates barplots using ggplot2.

```{r q4_ibu_medians, fig.height= 10}
#Get Medians
state_ibu_medians <- tapply(final$IBU, final$State, median, na.rm = T)
state_ibu_medians <- data.frame(state_ibu_medians)

#Reset Column Names
setDT(state_ibu_medians , keep.rownames = TRUE)
colnames(state_ibu_medians) <- c("State", "IBU_Medians")

#Bar Plot
p <- ggplot(data = state_ibu_medians, aes(x=State, y=IBU_Medians, width = .5)) +
    labs(title="IBU", x="State", y = "IBU") 

p + geom_bar(stat="identity") +
    theme(plot.title = element_text(hjust = 0.5), legend.position = "none") +
    coord_flip() +
    theme_minimal()
```


```{r q4_abv_medians, fig.height= 10}

#ABV Medians
state_abv_medians <- tapply(final$ABV, final$State, median, na.rm = T)
state_abv_medians <- data.frame(state_abv_medians)

#Reset Column Names
setDT(state_abv_medians , keep.rownames = TRUE)
colnames(state_abv_medians) <- c("State", "ABV_Medians")

#Bar Plot
p <- ggplot(data = state_abv_medians, aes(x=State, y=ABV_Medians, width = .5)) +
    labs(title="ABV", x="State", y = "ABV") 

p + geom_bar(stat="identity") +
    theme(plot.title = element_text(hjust = 0.5), legend.position = "none") +
    coord_flip() +
    theme_minimal()
```


<br/>

#### Question 5
Given that our client\'s beer is very strong and bitter, we would like to study markets with these same specs. The code below returns the states with the highest ABV and IBU.
```{r q5}
#State with max ABV
final[which.max(final$ABV), "State"]

#State with max IBU
final[which.max(final$IBU), "State"]
```

<br/>  

#### Question 6
```{r q6}
summary(final$ABV)
```

<br/>

#### Question 7
```{r q7}
p <- ggplot(final, aes(x=ABV, y=IBU)) + geom_point(shape=16) +  geom_smooth(method=lm) 
p + theme_hc()+ scale_colour_hc()
```