---
title: "CODEBOOK"
author: "unique(TeamName)"
date: "December 26, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(plyr)
```

# Codebook: Hoppy Specter Brewery Analysis


### Data Sets

The Hoppy Specter Analysis relies on two files contained in the <data> directory. These include: 

* Beers.csv 
  + Contains information about individual beers. 
* Breweries.csv
  + Contains informationa bout breweries producing beers. 
  
```{r}
#Read in the beers and breweries files
  Beers = read.csv("data/Beers.csv", stringsAsFactors = FALSE)
  Breweries = read.csv("data/Breweries.csv", stringsAsFactors = FALSE)
```
  
### Variable Names: 

Initial variable names found in the data sets are: 

#### Beers.csv Variables: 

1. Name -- Specifies the Beer Name

2. Beer_ID -- A unique identifier for the individual beer

3. ABV -- Alcohol By Volume -- The value of alcohol in a beer by volume. Multiplying this number by 100% gives percent alcohol by volume. 
  + If NA the value is missing. 

4. IBU -- International Bitterness Units -- A numerical score used to evaluate the bitterness of individual beers. 
  + If NA the value is missing. 
  
5. Brewery_id -- Unique identifier for the brewery that makes the beer. 
  
6. Style -- Presents the style of beer according to customary standards

7. Ounces -- Default packaging size

```{r}
  names(Beers)
```

#### Breweries.csv Variables: 

1. Brew_id -- Unique identifier for the brewery

2. Name -- Specifies the name of the brewery

3. City -- Specifies the City in which the brewery is located

4. State -- Specifies the Abbreviation of the State in which the brewery is located

```{r}
  names(Breweries)
```

#### Data preparation: 

The State records in the Breweries data have undesired white space in each record. We will remove that here. 

```{r}
Breweries$State = trimws(Breweries$State)
```


#### Combined Variable Table

In order to utilize these variables, we will normalize them and merge the data set here. 

```{r}
# Set a common Brewery ID variable to merge on: 

names(Breweries)[1] = "Brewery_id"

# Merge the two dataframes

MergedBrewing = merge(Beers,Breweries, by = "Brewery_id")

# Fix the names conflict

names(MergedBrewing)[2] = "Beer_Name"
names(MergedBrewing)[8] = "Brewery_Name"
head(MergedBrewing)
```

The final data frame contains the variables and types as follows: 

```{r}
variableName = names(MergedBrewing)
types = sapply(MergedBrewing,class)
variables = rbind (variableName,types)
knitr::kable(variables)
```

####Additional Data Frames Used / Created

One data frame is created to show the number of breweries in each state: 

```{r}

brew_count <- count(Breweries$State)
colnames(brew_count) <- c("State", "Brewery_Count")
brew_count <- brew_count[with(brew_count,order(-Brewery_Count)), ]
knitr::kable(brew_count)
```

It has types: 

```{r}

variableName = names(brew_count)
types = sapply(brew_count,class)
variables = rbind (variableName,types)
knitr::kable(variables)
```

Two additional data frames are used to identify and track the ABV and IBU Median values per state. 

```{r}

# Obtain ABV Data: 

#ABV Medians
state_abv_medians <- tapply(MergedBrewing$ABV, MergedBrewing$State, median, na.rm = T)
state_abv_medians <- data.frame(state_abv_medians)

#Reset Column Names
setDT(state_abv_medians , keep.rownames = TRUE)
colnames(state_abv_medians) <- c("State", "ABV_Medians")

head(state_abv_medians)

# Obtain IBU Data: 

#Get Medians
state_ibu_medians <- tapply(MergedBrewing$IBU, MergedBrewing$State, median, na.rm = T)
state_ibu_medians <- data.frame(state_ibu_medians)
state_ibu_medians <- na.omit(state_ibu_medians)

#Reset Column Names
setDT(state_ibu_medians , keep.rownames = TRUE)
colnames(state_ibu_medians) <- c("State", "IBU_Medians")

head(state_ibu_medians)
```

The data types in these tables are: 

```{r}
variableName = names(state_ibu_medians)
types = sapply(state_ibu_medians,class)
variables = rbind (variableName,types)
knitr::kable(variables)

variableName = names(state_abv_medians)
types = sapply(state_abv_medians, class)
variables = rbind (variableName,types)
knitr::kable(variables)
```